---
title: "SQL"
output: html_document
---

#Filtering Columns and Rows
By installing Lahman and sqldf you can filter columns and rows with information from different data sets. It is useful for organization. 
```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)
```

```{r}
query<- "SELECT playerID, yearID, teamID, HR
FROM Batting 
WHERE teamID='NYA' AND yearID=1927"
sqldf(query)
```

```{r}
query<- "SELECT playerID, yearID, teamID, HR
FROM Batting
WHERE HR>=40 and teamID='NYA'"
sqldf(query)
```

```{r}
query<- "SELECT playerID, teamID, HR, SO, yearID
FROM Batting 
WHERE HR>40 and SO<60"
sqldf(query)
```

```{r}
query<- "SELECT teamID, yearID, playerID, HR
FROM Batting
WHERE teamID='PHI' and yearID >=1970 and yearID <1980 and HR>30"
sqldf(query)
```

# Order By
By using the Order By function, you can take the information you retrieved from the dataset and organize them by their individual rows.

```{r}
query<- "SELECT playerID, HR, teamID, yearID
FROM Batting 
WHERE HR>50 
ORDER BY HR DESC"
sqldf(query)
```

```{r}
query<- "SELECT playerID, teamID, yearID, HR, SO, AB
FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

#AGGREGATION 
Aggregation is producing the minimum amount of output.

```{r}
query<- "SELECT playerID, sum(HR) 
FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

```{r}
query<- "SELECT playerID, avg(HR)
FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

# JOINS
You can tables from multiple data tables using joins

```{r}
query<- "SELECT nameFirst, nameLast, teamID, yearID, HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

```{r}
query<- "SELECT nameFirst, nameLast, teamID, yearID, HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

```{r}
query<- "SELECT playerID, name, Batting.yearID, Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

```{r}
query<-"SELECT nameFirst,nameLast,sum(HR)
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```

